Exploring probability distributions for bivariate temporal granularities

### 2019-12-09 main.Rmd
till quasi-circular

\begin{equation} \label{eq:eq4}
\begin{split}
Q_{B,G'}(8) & = L(8 - \sum_{w=0}^{3-1}\vert T_{w \mod 2}\vert) , 8 \in T_{3} \\
  & = L(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\
  & = L(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\
  & = L(8 - \vert T_{0 \mod 2}\vert - \vert T_{1 \mod 2}\vert - \vert T_{2 \mod 2}\vert) \\
  & = L(8 - 2*\vert T_{0 \mod 2}\vert - \vert T_{1 \mod 2}\vert) \\
  & = L(8 - 2*2 - 3) \\
  & = L(1) \\
\end{split}
\end{equation}

\begin{equation} \label{eq:eq5}
\begin{split}
Q_{B,H'}(10) & = L(10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert) ,  10 \in T_{1}  \\
  & = L(10 - \sum_{w=0}^{0}\vert T_{w\mod 2}\vert) \\
  & = L(10 - \vert T_{0 \mod 2}\vert) \\
  & = L(10 - \vert T_{0}\vert) \\
  & = L(10 - 7) \\
  & = L(3) \\
\end{split}
\end{equation}

@@

Here $w \mod R'$ represents the index that must be shifted to obtain $G'(w)$. The idea here is if we know the composition of each of the granules of $G'$ in terms of granules of $B$ for one period, we can find the composition of any granule of $G'$ beyond a period since the "pattern" repeats itself along the time domain due to the periodic property. The periodic property also ensures that $\vert T_w \vert = \vert T_{w\mod R'} \vert$ since every $w^{th}$ and $(w+R)^{th}$ granule of $G'$ will have the same number of granules of $B$. The term $\sum_{w=0}^{k-1}\vert T_{w}\vert$ denotes the number of granules of $B$ till the $(k-1)^{th}$ granule of $G'$. Since $\vert T_w \vert = \vert T_{w\mod R'} \vert$, the number of granules of $B$ till the $(k-1)^{th}$ granule of $G'$ becomes $\sum_{w=0}^{k-1}\vert T_{w \mod R'}\vert$ in in \autoref{def:quasicircular}.

@@

It can be noted here that if a linear granularity $G'$ is quasi-periodic with respect to $B$, then \autoref{eq:quasi} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$ (by Definition \autoref{def:quasi}).

@@

Aperiodic cyclic granularities are defined using aperiodic linear granularities. Consider n aperiodic linear granularities $M_i$  $\forall \{i\in{1, 2, \dots, n\}}$ none of which can be expressed as periodic or quasi-periodic with respect to the bottom granularity. Further assume that $B \trianglelefteq M_i$ $\forall \{i\in{1, 2, \dots, n\}}$. Then according to \autoref{def:groupsinto}, for each index $j$ there exists a (possibly infinite) subset $T_{\{i_j\}}$ of the integers such that
\begin{equation*}
M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)
\end{equation*}
and suppose $M = \bigcup_{i=1}^{n}M_{i}$ is formed by collecting the granules of $\{M_1, M_2, \dots, M_n\}$.

\begin{definition}\label{aperiodic}
An aperiodic cyclic granularity $A_{B, M}$ relates aperiodic linear granularity $M$ and bottom granularity $B$, if

\begin{equation}\label{eq:aperiodic}
\begin{split}
A_{B, M}(z) & = L(i), \quad z \in T_{i_j} \quad and \quad \forall i \in {1, 2, \dots, n}\\
     & = L(0), \quad otherwise
\end{split}     
\end{equation}
where, 
$z \in \mathbb{Z}_{\geq 0}$ denotes the index set,
$T_{i_j}$ are the sets of indices of $B$ describing $M_{i}(j)$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$ and $M = \bigcup_{j=1}^{n}M_{j}$.
\end{definition}

@@

**Discussion**: The number of categories obtained coincides with the number of aperiodic linear granularities considered and an additional category representing no event. The index ${\{i_j\}}$ stands for the $j^{th}$ granule of the $i^{th}$ linear aperiodic granularity considered.

@@

In \autoref{sec:cyclic-gran}, we discussed how we can define cyclic granularities by extending the concept of linear granularities defined in @Bettini1998-ed. In this section, we will see how to obtain cyclic granularities through an algebraic representation of other cyclic granularities. This is similar to calendar algebra in @Ning2002-tf, where linear granularities are generated from the bottom linear granularity or from those that are already generated. Since, our method caters to computation of cyclic granularities, we shall refer to it as "cyclic calendar algebra".

>>>>

\begin{equation} \label{eq:eq4}
\begin{split}
Q_{B,G'}(8) & = L(8 - \sum_{w=0}^{3-1}\vert T_{w \mod 2}\vert) ,\quad since \quad 8 \in T_{3} \\
  & = L(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\
  & = L(8 - \sum_{w=0}^{2}\vert T_{w\mod 2}\vert) \\
  & = L(8 - \vert T_{0 \mod 2}\vert - \vert T_{1 \mod 2}\vert - \vert T_{2 \mod 2}\vert) \\
  & = L(8 - 2*\vert T_{0 \mod 2}\vert - \vert T_{1 \mod 2}\vert) \\
  & = L(8 - 2*2 - 3) \\
  & = L(1) \\
\end{split}
\end{equation}

\begin{equation} \label{eq:eq5}
\begin{split}
Q_{B,H'}(10) & = L(10 - \sum_{w=0}^{1-1}\vert T_{w \mod 2}\vert) ,\quad since \quad  10 \in T_{1}  \\
  & = L(10 - \sum_{w=0}^{0}\vert T_{w\mod 2}\vert) \\
  & = L(10 - \vert T_{0 \mod 2}\vert) \\
  & = L(10 - \vert T_{0}\vert) \\
  & = L(10 - 7) \\
  & = L(3) \\
\end{split}
\end{equation}

@@

Here $w \mod R'$ represents the index that must be shifted to obtain $G'(w)$. The idea here is if we know the composition of each of the granules of $G'$ in terms of granules of $B$ for one period, we can find the composition of any granule of $G'$ beyond a period since the "pattern" repeats itself along the time domain due to the periodic property. The periodic property also ensures that $\vert T_w \vert = \vert T_{w\mod R'} \vert$ since every $w^{th}$ and $(w+R)^{th}$ granule of $G'$ will have the same number of granules of $B$. The term $\sum_{w=0}^{k-1}\vert T_{w}\vert$ denotes the number of granules of $B$ till the $(k-1)^{th}$ granule of $G'$. Since $\vert T_w \vert = \vert T_{w\mod R'} \vert$, the number of granules of $B$ till the $(k-1)^{th}$ granule of $G'$ becomes $\sum_{w=0}^{k-1}\vert T_{w \mod R'}\vert$ in Definition \autoref{def:quasicircular}.

@@

It can be noted here that if a linear granularity $G'$ is quasi-periodic (Definition \autoref{def:quasi}) with respect to $B$, then \autoref{eq:quasi} can be modified as follows to account for exceptions $E = [e_{begin}, e_{end}]$ (by Definition \autoref{def:quasi}).

@@

Aperiodic cyclic granularities are defined using aperiodic linear granularities. Consider n aperiodic linear granularities $M_i$  $\forall \{i\in{1, 2, \dots, n\}}$ (that is, none of them can be expressed as periodic or quasi-periodic with respect to the bottom granularity). Further assume that $B \trianglelefteq M_i$ $\forall \{i\in{1, 2, \dots, n\}}$. Then according to \autoref{def:groupsinto}, for each index $j$ there exists a (possibly infinite) subset $T_{\{i_j\}}$ of the integers such that
\begin{equation*}
M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)
\end{equation*}
and suppose $M = \bigcup_{i=1}^{n}M_{i}$ is formed by collecting the granules of $\{M_1, M_2, \dots, M_n\}$.

\begin{definition}\label{aperiodic}
(aperiodic cyclic granularity): An aperiodic cyclic granularity $A_{B, M}$ relates aperiodic linear granularity $M$ and bottom granularity $B$, if

\begin{equation}\label{eq:aperiodic}
\begin{split}
A_{B, M}(z) & = L(i), \quad z \in T_{i_j} \quad and \quad \forall i \in {1, 2, \dots, n}\\
     & = L(0), \quad otherwise
\end{split}     
\end{equation}
where, 
$z \in \mathbb{Z}_{\geq 0}$ denotes the index set,
$T_{i_j}$ are the sets of indices of $B$ describing aperiodic linear granularities $M_{i}(j)$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$, $M = \bigcup_{j=1}^{n}M_{j}$ and index ${\{i_j\}}$ stands for the $j^{th}$ granule of the $i^{th}$ linear aperiodic granularity considered.
\end{definition}

@@

**Discussion**: The number of categories obtained coincides with the number of aperiodic linear granularities considered and an additional category representing no event.

@@

In \autoref{sec:cyclic-gran}, we discussed how we can define cyclic granularities by extending the framework of linear granularities defined in @Bettini1998-ed. In this section, we will see how to obtain cyclic granularities through an algebraic representation of other cyclic granularities. This is similar to calendar algebra in @Ning2002-tf, where linear granularities are generated from the bottom linear granularity or from those that are already generated. Since, our method caters to computation of cyclic granularities, we shall refer to it as "cyclic calendar algebra".
